Type-Safe Modular Parsing

### 2017-06-05 SLE_17/paper.tex
More.
A bit.
Fit into 12 pages
Revised abstract and intro

%% For double-blind review submission
\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=false}
%% For single-blind review submission
%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[sigplan,10pt]{acmart}\settopmatter{}

@@

@@

%% Paper note
%% The \thanks command may be used to create a "paper note" ---
%% similar to a title note or an author note, but not explicitly
%% associated with a particular element.  It will appear immediately
%% above the permission/copyright statement.
\thanks{with paper note}                %% \thanks is optional
                                        %% can be repeated if necesary
                                        %% contents suppressed with 'anonymous'

@@

  This paper investigates solutions for the problem of \emph{modular
    parsing}. We focus on \emph{semantic} modularity and not just
  \emph{syntactic} modularity. That is, the solutions should not only
  allow complete parsers to be built out of modular parsing
  components, but also enable the parsing components to be
  \emph{modularly type-checked} and \emph{separately compiled}. We
  present a technique based on parser combinators that enables modular
  parsing. Interestingly, the modularity requirements for modular
  parsing rule out several existing parser combinator approaches,
  which rely on some non-modular techniques. We show that Packrat
  parsing techniques, provide solutions for such modularity problems,
  and enable reasonable performance in a modular setting.
  Extensibility is achieved using \emph{multiple inheritance} and
  \emph{Object Algebras}.  To evaluate
  the approach we conduct a case study based on the ``Types and
  Programming Languages'' interpreters. The case study shows the
  effectiveness at reusing parsing code from existing interpreters,
  and the total parsing code is 69\% shorter than an existing code
  base using a non-modular parsing approach.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional

@@

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

>>>>

%% For double-blind review submission
\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=false}
\settopmatter{printacmref=false}
%% For single-blind review submission
%\documentclass[sigplan,10pt,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[sigplan,10pt]{acmart}\settopmatter{}

@@

\setlength{\abovecaptionskip}{1.5mm}
\setlength{\belowcaptionskip}{1.5mm}

\setlength{\textfloatsep}{2mm}

@@

%% Paper note
%% The \thanks command may be used to create a "paper note" ---
%% similar to a title note or an author note, but not explicitly
%% associated with a particular element.  It will appear immediately
%% above the permission/copyright statement.
%%\thanks{with paper note}                %% \thanks is optional
                                        %% can be repeated if necesary
                                        %% contents suppressed with 'anonymous'

@@

  This paper investigates solutions for the problem of \emph{modular
    parsing}. We focus on \emph{semantic} modularity and not just
  \emph{syntactic} modularity. That is, the solutions should not only
  allow complete parsers to be built out of modular parsing
  components, but also enable the parsing components to be
  \emph{modularly type-checked} and \emph{separately compiled}. We
  present a technique based on parser combinators that enables modular
  parsing. 
  %Interestingly, the modularity requirements for modular
  %parsing rule out several existing parser combinator approaches,
  %which rely on some non-modular techniques. 
  We show that Packrat
  parsing techniques, provide solutions for such modularity problems,
  and enable reasonable performance in a modular setting.
  Extensibility is achieved using \emph{multiple inheritance} and
  \emph{Object Algebras}.  To evaluate
  the approach we conduct a case study based on the ``Types and
  Programming Languages'' interpreters. The case study shows the
  effectiveness at reusing parsing code from existing interpreters,
  and the total parsing code is 69\% shorter than an existing code
  base using a non-modular parsing approach.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
%\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional

@@

%% Appendix
%% \appendix
%% \section{Appendix}

%% Text of appendix \ldots

### 2017-06-05 SLE_17/sections/Introduction.tex
Paper
Revised abstract and intro

To address the modularity limitations of Programming Languages, several
different approaches have been proposed in the past. Existing
approaches can be broadly divided into two categories:
\emph{syntactic} or \emph{semantic} modularization
techniques. Syntactic modularization techniques are quite popular in
practice, due to their simplicity of implementation and use.
Examples include many tools for developing Feature-Oriented Software-Product
Lines (SPLs)~\cite{AK:JOT09,Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.  Most syntactic approaches employ textual
composition techniques such as \emph{superimposition}~\cite{AK:JOT09} to
enable the development modular program features. Such
composition techniques collect the code for multiple features and
merge them together when a concrete combination of features is needed
for a particular program. As Kastner et. al~\cite{Kastner11road} note,
a typical drawback of feature-oriented SPL implementations, which
more generally applies to syntactic modularity approaches, is that
such ``\emph{implementation mechanisms lack proper
  interfaces and support neither modular type checking nor separate
  compilation}''.

@@

\item \textbf{TAPL case study:} We conduct a case study with 18 interpreters
  from the TAPL book. The case study show the effectiveness of modular
  parsing in terms of reuse. The TAPL case study is available online
  at:
  \url{http://}\footnote{\textbf{Note to Reviewers:} Please find the URL in the
    supplementary material for the submission.}

@@

All the code in the paper and the case study is written in Scala, since its concise and elegant
syntax gives better presentation. Clearly
other languages that support some form of multiple inheritance
(including Java 8 with default methods~\cite{Goetz2012}) could in principle be
used.

>>>>

To address the modularity limitations of Programming Languages, several
different approaches have been proposed in the past. Existing
approaches can be broadly divided into two categories:
\emph{syntactic} or \emph{semantic} modularization
techniques. Syntactic modularization techniques are quite popular in
practice, due to their simplicity of implementation and use.
Examples include many tools for developing Feature-Oriented Software-Product
Lines (SPLs)~\cite{AK:JOT09,Kastner11road}, some Language Workbenches~\cite{Erdweg201524}, or extensible parser
generators~\cite{antlr1995,Grimm2006,Gouseti2014,Warth2016}.  Most syntactic approaches employ textual
composition techniques such as \emph{superimposition}~\cite{AK:JOT09} to
enable the development modular program features. 
%Such
%composition techniques collect the code for multiple features and
%merge them together when a concrete combination of features is needed
%for a particular program. 
As Kastner et. al~\cite{Kastner11road} note,
a typical drawback of feature-oriented SPL implementations, which
more generally applies to syntactic modularity approaches, is that
such ``\emph{implementation mechanisms lack proper
  interfaces and support neither modular type checking nor separate
  compilation}''.

@@

\item \textbf{TAPL case study:} We conduct a case study with 18 interpreters
  from the TAPL book. The case study show the effectiveness of modular
  parsing in terms of reuse. The TAPL case study is available online
  at:
  \url{http://}\bruno{Please add the link here!}

@@

All the code in the paper and the case study is written in Scala, since its concise and elegant
syntax gives better presentation. Other languages that support some form of multiple inheritance
(including C++ or Java 8 with default methods~\cite{Goetz2012}) could in principle be
used.
